---
title: "New Wooldridge Package!"
output: html_notebook
---

```{r setup}
knitr::opts_chunk$set(message=FALSE, warning=FALSE)
install.packages("tidyverse")
install.packages("wooldridge")

library(tidyverse)
library(wooldridge)
```

Attention econ students, professors and aficianados - an awesome new R package has arrived for the fall semester. It's called `wooldridge` and as you might expect, it's a companion R package to the ~~Bible of econometrics~~ popular Wooldridge text used in lots of econometrics classes. Thanks to [@justinmshea](https://www.linkedin.com/in/justinmshea/) for building and contributing to CRAN!

The [vignette](https://cran.r-project.org/web/packages/wooldridge/vignettes/wooldridge-vignette.pdf) has a nice summary and worked example from every chapter of the book - here's an excerpt:  

> This vignette contains examples from every chapter of *Introductory Econometrics: A Modern Approach* by Jeffrey M. Wooldridge. Each example illustrates how to load data, build econometric models, and compute estimates with R. Economics students new to both econometrics and R may find the introduction to both a bit challenging. In particular, the process of loading and preparing data prior to building one’s first econometric model can present challenges. The wooldridge data package aims to lighten this task. 

Honestly, the best thing to do is head straight to the vignette but below is a quick worked example from Chapter 10 on time series. 

```{r}
# Load up the data for Chapter 10, example 10.2 and take a look at the summary
data("intdef")
summary(intdef)
```

Now let's run example 10.2 and examine the effects of inflation and deficits on interest rates.

The variable *i3* is the three-month Treasury-bill rate, *inf* is the annual inflation rate based on the consumer price index (CPI), and *def* is the federal budget deficit as a percentage of GDP. The  equation to be estimated is is:

$$\hat{i3_{t}}=\beta_0 + \beta_1inf_{t} + \beta_2def_{t}  + e$$
We will run the same regression as the book and the vignette. The only wrinkle is that we will use the `broom` package to the clean up the results and visualize predicted values.

```{r}
# Run regression
tbill_model <- lm(i3 ~ inf + def, data = intdef)

# tidy the results
tidy(tbill_model)
```

We can `glance` at our results, and use dplyr's `select` verb to choose a handful of columns for viewing. 

```{r}
glance(tbill_model) %>% 
  select( r.squared, adj.r.squared, sigma)
```


Let's round out our use of `broom` and tinker with the `augment` function, which will augment our original data set with fitted/predicted values and residuals from the model.

```{r}
intdef_augmented <- augment(tbill_model, intdef)

head(intdef_augmented)
```

Now we can visualize the predicted or `.fitted` versus actual *i3* values. 

```{r}
intdef_augmented %>% 
  ggplot(aes(x = year)) + 
  geom_line(aes(y = i3, color = "i3")) +
  geom_line(aes(y = .fitted, color = "predicted"))

```


Let's add in our predictors as well and see if anything jumps out as interesting. 

```{r}
intdef_augmented %>% 
  ggplot(aes(x = year)) + 
  geom_line(aes(y = i3, color = "i3")) +
  geom_line(aes(y = .fitted, color = "predicted")) +
  geom_line(aes(y = inf, color = "inflation")) +
  geom_line(aes(y = def, color = "deficit"))
```


Since 2000, interest rates have been heading in one direction, while the deficit as percent of GDP has been increasing. Remember that our model returned a positive beta for the `def` variable: an increasing deficit should lead to increasing interest rates. This might explain why the predictions are way off after 2000.

That's all for today. Thanks again to [justinmshea](https://www.linkedin.com/in/justinmshea/) for the new fantastically useful [wooldridge](https://cran.r-project.org/web/packages/wooldridge/wooldridge.pdf) package. Happy econometricsing! 